\setlength{\parindent}{0pt}
\section{The Story of Babel\idx{Babel} -- {\small Jonathan Roose}}
{
\def\drie{\vspace{0.9cm}}
\renewcommand{\glyphkernright}{-0.25em}
\renewcommand{\glyphkernleft}{-0.15em}
\catcode`\~=13\def\~{\hspace{0.4cm}}
\baselineskip=1.2\baselineskip

%\def\isyo#1{\scalerel*{#1}{24pt}}

%Dit is een test \isyo{\ep} dit is een test. 


\catcode`\<=13\def<{\kern-.2em}

\at\kern-.2em\ej\on	~	\af~	\at<\tem		~\pa\tu\ta	 

{\it AT.EJ.ON		AF	AT.TEM		PA.TU.TA	 }

{\tt ALL.PERSON.PLURAL   	ON	ALL.PL-EARTH	PAST.BE	 }

Everybody		on 	Earth 		had

\drie

\ek\ip\lan\Atlansum~\an~\ek\set\on\me\comma

{\it EK.IP.(A).LAN.SUM	AN	EK.SET.ON.ME.}

{\tt ACC.ONE.SAY.COMU.	AND.	ACC.WORD.PLURAL.SAME}

the same language	 and 	the same words. 

\drie


And	 		it came to pass, 			  

{\tt REL-CLAUSE.AND            AT.TIME.THIS.PAST.HAPPEN 		}

{\it I.AN			ET.JA.ES.PA.NES	}

\Atlani\an                   ~ \et\ja\es\pa\nes

\drie

\Atlani\po\ja	~\aj\on 

{\it I.JA.PO.	EJ.AJ.ON }

{\tt REL-CLAUSE.PROGR.TIME	THR.PERS.PLURAL }

as	they	    

\drie 

\pa\tu\mef\lu\kem\ ~ \li\jol

{\it PA.TU.MEF.LU.KEM LI.JOL }

{\tt VERB.MOVE.PLACE.HOME	DIRECTION.ORIGIN	 }

migrated          			from      		 

\drie 

\lu\pes\son ~  \aj\on

{\it LU.PES.SON 	EJ.AJ.ON }

{\tt PLACE.BIRTH.SON	PERSON.3RD-REMOVED.PLURAL }

the east, 		they 		 


\drie 


\Atlanni\faf~\ek\pas\mi\man

{\it NI.FAF~                                EK.PAS.MI.MAN }

{\tt PERFECTED.ECOUNTER  	ACC.PLAIN.BETWEEN. MOUNTAIN.	}

came upon 		a valley                 


\drie  


\et\mes~\lu\na\cartouche{\Atlans\ji\nal}\period

{\it ET.MES~		LU.NA.[S.JI.NAL]. }

{\tt AT.COUNTRY	PLACE.NAME.[SHINAR] }

in the land          of Shinar. 

\drie



\an~\aj\on~\pa\tu\men\Atlansup\ken

{\it AN~	AJ.ON~   PA.TU.MEN.SUP.KEN }

{\tt AND.	3TH.PLURAL	PAST.VERB.MAKE.SIT.HOME	 }

and 	they 		settled			 

\drie
  


\es\lu\aj~\period \aj\on~\pa\lan

{\it ES.LU.AJ~		AJ.ON~		PA.LAN	 }

{\tt DEMONST.2ND.PLACE	3TH.PLURAL	PAST.SAY }

there.			They		said 	 

\drie 


\lo\su\of\on\comma

{\it LO.SU.OF.ON, }

{\tt DATI.SELF.OTHER.PLURAL		 }
 
to one another, 			 
\drie 



  

\Atlano\comma\tu\men\am\on\

{\it O,TU.MEN.AM.ON }

{\tt IMPERATIVE.VERB.MAKE.1ST.REMOVED.PLURAL }

“come, let us make   


\drie 


\ek\fuk\jet\on\comma\Atlani\an\

{\it EK.FUK.JET.ON~			I.AN }

{\tt ACC.BUILD-BLOCK.STONE.PLU 	REL.CAUSE.AND		 }

bricks			,and 	 

\drie
  

\ek\es~\tu\pen\pin\toj\period

{\it EK.ES TU.PEN.PIN.TOJ.}

{\tt VERB.BURN.CAUSE.BECOME.SOLID}

burn them			hard.” 
\drie


\fuk\jet\on ~ \tu\pa\si

{\it FUK.JET.ON			TU.PA.SI			 }

{\tt BUILD-BLOCK.STONE.PLUR          VERB.PAST.PREDICATE	 }

Bricks 				served 				 
\drie

\lo\aj\on ~ \me ~\fuk\maj\jet

{\it LO.AJ.ON		ME	FUK.MAJ.JET }

{\tt DAT.3TH-PERSON.PLUR AS	BUILDING.MATTER.STONE. }

them			as 	stone\idx{tone}	 


\drie
  

\an ~ \pat\na\cartouche{\Atlanpi\tu\mum} ~ \tu\pa\si

{\it I.AN	PAT.NA[BI.TU.MUM]	TU.PA.SI }

{\tt REL.CLAUSE.AND	NAMED[BITUMUM]	TU.PAST.PREDICATE }

and	bitumum 		served  
\drie


  

\lo\aj\on ~\me ~\fuk\maj\mop\period

{\it LO.AJ.ON			ME	FUK.MAJ.MOP. }

{\tt DATI.3TH-PERSON.PLUR               AS	BUILDING.MATTER.FOAM }

them				as 	mortar. 

\drie
  

\Atlani\an ~ \ej\aj\on ~ \pa\lan

{\it I.AN			EJ.AJ.ON			PA.LAN }

{\tt RELA-CLAUSE.AND          PERSONS.3TH-REMO.PLURAL 	PAST.SAY }

And 			they 				said  

\drie

  

\Atlano\tu\men\nap\am\on

{\it O.TU.MEN.NAP.AM.ON }

{\tt IMPERATIVE.VERB.MAKE. BUILDING.1ST.REMOVED.PLURAL	 }

“come, let us build                                                                          
\drie
 

\ek\ip\tos\comma~\an~\ek\nap\lak\te

{\it EK.IP.TOS		,AN	EK.NAP.LAK.TE	 }

{\tt ACC.ONE.TOWN		AND	ACC.BUILDING.LONG.VERTICAL }

a city			,and 	a tower	 
\drie

\ta\es\aj\fon\Atlanin\som

{\it TA.ES.AJ.FON.IN.SOM }

{\tt GENA.THIS.3TH-REMO.PEAK	IN	SKY }

with its top 				in           the sky 
\drie

  

\tu\men~	\ek\na~		\lo\am\on 

{\it TU.MEN	EK.NA		LO.AM.ON }

{\tt VERB.MAKE	ACC.NA		DAT.1ST-REMO.PLURAL	 }

To make 	a name 		for ourselves		 

\drie

 

\Atlani\of\Atlanpi ~ \am\on

{\it I.OF.PI			AM.ON }

{\tt REL.CLAUSE.OTHER	1ST-REMO.PLURAL }

,else 			we 

\drie
 

\po\si~\tu\mej\Atlanne\ip

{\it PO.SI			  TU.MEJ.NE.IP		 }

{\tt FUTURE.PROGRESSIVE.PREDICATE    VERB.CHANGE.NEGATION.ONE        }

shall be			    scattered			
\drie 


\ki\an\fa ~ \lok\tem\period

{\it KI.AN.FA		LOK.TEM. }

{\tt NEAR.AND.FAR		ROUND.PLAN-EARTH }

all over			The earth.
\drie 
 
\Atlana\jel ~ \tu\pa\kom\lit 

{\it A.JEL			TU.PA.KOM.LIT			 }

{\tt STRESS.GOD         VERB.PAST.COME.DOWN		}

the LORD 		came down 			

\drie 

\fan\tu\sik ~ \ek\tos

{\it FAN.TU.SIK		EK.TOS }

{\tt DESIRE.VERB.SEE	ACC.CITY }

to look			at the city 

\drie 
  
\an~\ek\nap\lak\te ~\es

{\it AN	EK.NAP.LAK.TE			ES			 }

{\tt AND	ACC.BUILDING.LONG.VERTICAL	DEMONSTRATIVE	 }

and 	tower 				that 			 

\drie 

\ej\Atlansum\on ~ \po\ken\nap

{\it EJ.SUM.ON			PO.KEN.NAP }

{\tt COMMUNITY.PERSON.PLURAL	PERFECTED.MAKE.BUILDING }

men				had built 

  
\drie 

\Atlani\an ~ \Atlana\jel~\pa\lan

{\it I.AN			A.JEL		PA.LAN }

{\tt REL.CLAUSE.AND              STRESS.GOD	PAST.SAY }

and 			the LORD            said 

\drie 

\Atlani\Atlanif~\pe\at\ip\Atlansum

{\it I.IF			PE.AT.IP.SUM				 }

{\tt REL.ClAUSE.CONDITION	POSSIBLE.ALL.ONE.COMMUNITY		 }

“if, 			as one people 		
\drie 


\ut\ip\lan\Atlansum~\ta\aj\at\ej\comma

{\it UT.IP.LAN.SUM			TA.AJ.AT.EJ }

{\tt INSTRU.ONE.SAY.COMMUNITY	GENA.3TH-REMO.ALL.PERSON	 }

with one language 		for all, 
\drie 

\am\es\si~\tu\ut~\aj\on
 
{\it AM.ES.SI			TU.UT		AJ.ON }

{\tt 1ST-RELA.DEMONSTR.PRED.	VERB.INSTRU.    3TH-REMO.PLURAL }

this is 				how       	they  
  
\drie 

\Atlanni\ka~\tu\kuk~\is\ku

{\it NI.KA				TU.KUK		IS.KU                    } 

{\tt PERFECTED.BEGINNING.                 VERB.CUSTOM	CONC.CONTR	 }

have begun 			to act		then 		 
\drie 


\is\ol~\es~\aj\on

{\it IS.OL		ES		AJ.ON	 }

{\tt PREDAC.OR	DEMONSTRA.	3TH-REMO.PLURAL }

nothing 	that		They 	 

\drie 

 
\tu\lan\Atlanpi\Atlannil~\Atlanfi\tu~\Atlanfi\si

{\it TU.LAN.PI.NIL		               FI.TU		FI.SI }

{\tt VERB.SAY.POSSIBL.IMAGI	FUTR.VERB	FUTR.PREDICATE }

may propose			to do		will be 
\drie 

  

\ap~\ta\aj\on~\Atlanmu\Atlanto\jam\period

{\it AP		TA.AJ.ON			MU.TO.JAM. }

{\tt OUTSIDE	GENI.3TH-REMO.PLURAL	META.RANGE.HAND }

out 		of their 				reach.
\drie

  
\am\on~\is~\Atlano\uf\mef\li\ot

{\it AM.ON,     IS,   O-UF.MEF.LI.OT				 }

{\tt 1ST-REMO.PLURAL CONCLU EXCLA.INTENT.GO.TOWAR.DOWN             }

let us, then, go down, 			      	 

\drie
 

\Atlani\an~\tu\lis\nef\Atlansin

I.AN			TU.LIS.NEF.SIN      

{\tt RELA-CLAU.AND	VERB.DISSOCIATION.KNOW.THINK }

and  			confound\idx{sound} 

\drie

\ek\ta\aj\on~\ek\me\set\on

{\it EK.TA.AJ.ON		EK.ME.SET.ON }

{\tt ACC.GEN.3TH.PLURAL             ACC.SAME.WORD.PLURAL	 	}

their 			speech

\drie

  

\es\aj\lu~\uf~\aj\on

{\it ES.AJ.LI				UF		AJ.ON}				 
{\tt DEMONSTR.3TH.PLACE        INTENT.		3TH.PLUR	 }


there			so 			they 		

\drie

\fe\Atlanne\nef\Atlansim~\su\ek\of\on

{\it FE.NE.NEF.SIM			SU.EK.OF.ON	 }

{\tt FUTURE.NEG.KNOW.THINK		ACC.SELF.OTHER.PLURAL }

shall not understand 		One another’s 

\drie

  
\ek\me\set\on\period

{\tt ACC.SAME.WORD.PLURAL }

{\it EK.ME.SET.ON }

speech.

\drie

 

\Atlani\is~\Atlano\jel 

{\it I.IS		O.JEL		 }

{\tt REL.CLAUSE.CONCLUSION.	EXCLA.GOD					 }

Thus 		the LORD 					
\drie

 

\tu\pa\si\Atlanne\ip~\ek\aj\on

{\it TU.PA.SI.NE.IP				EK.AJ.ON }

{\tt VERB.PAST.PREDICATE.NEGATION.ONE	ACC.3TH.PLUR }

scattered				them 
\drie

 

\li\jol~\es\aj\lu~\ki\an\fa

{\it LI.JOL			ES.AJ.LU		KI.AN.FA	 }

{\tt DIRECTION.ORIGIN         DEMONSTR.3TH.PLACE  NEAR.AND.FAR }

from 			there 			over 		
\drie
 

 
\at\sem\tem

{\it EK.AT.SEM.TEM }

{\tt ALL.SURFICE.PLAN-EARTH }

the face of the whole earth 


\drie

 

\an~\aj\on~\tu\Atlanne\mo\nes

{\it AN	AJ.ON		TU.NE.MO.NES		                               }

{\tt AND	3TH.PLUR	VERB.NEG.COMPA.TRANS.HAPPEN	 }

And 	they 		stopped 				 
\drie
 

\tu\ken\nap~\tos\period

{\it TU.KEN.NAP                      TOS.}

{\tt VERB.MAKE.BUILD	TOWN }

building 		the city. 
\drie

\ek\Atlana\tos~ \es\ko\si

{\it ES.KO.SI				EK.A.TOS		 }

{\tt DEMONSTR.REASEN.PREDICATE	ACC.STRESS.CITY	 }

That is why 				it 			

\drie
 
was called Babel\idx{Babel} 

{\it TU.PA.NA[PA.PEL] }

{\tt VERB.PAST.NAME[BABEL] }

\tu\pa\na\cartouche{\pa\pel}

\drie


\ko\fu~\es\aj\lu~\Atlana\jel

{\it KO.FU			ES.AJ.LU			A.JEL }

{\tt REASON.FIN-STATE          DEMOSTR.3TH.PLACE		STRESS.GOD }

Because 		there 				the LORD  
\drie

\tu\pa\lis\nef\Atlansin

{\it TU.PA.LIS.NEF.SIN}                                              			 
{\tt VERB.PAST.DISSOCIATION.KNOW.THINK	 }

confound\idx{sound}ed 
\drie

 

\at\tem\Atlansum~\ek\set\on\me\period

{\it EK.SET.ON.ME      AT.TEM.SUM.			}


{\tt ACC.SAME.WORD.PLURAL 	ALL.PLA-EARTH.COM. }

the speech			 of the whole earth.

\drie

\Atlani\an~\li\mi~\es\aj\lu

{\tt RELA-CLAU.AND	DERECTIO.ORIGIO      DEMONSTR.3TH.PLACE }

I.AN			LI.MI			ES.AJ.LU			 

And 			from 			there 		
\drie

\Atlana\jel~\tu\pa\si\Atlanne\ip

A.JEL		TU.PA.SI.NE.IP    

{\tt STRESS.GOD	VERB.PAST.PREDICATE.NEGATION.ONE }

the LORD	scattered 
\drie

 
\ek\aj\on~\ki\an\fa

{\it EK.AJ.ON	KI.AN.FA               }

{\tt ACC.3TH.PLUR	NEAR.AND.FAR	 }

them 		over 		 
\drie
  
\ek\at\sem\tem

{\it EK.AT.SEM.TEM }

{\tt ON.ALL.SURFACE.PLAN-EARTH }

the face of the whole earth. 
\drie
\vfill

\pagebreak
\section{The Epic of Gilgamesh\idx{Gilgamesh} -- {\small Niek Elsinga}}


\noindent {\bf Enkidu’s Dream - \Atlantan\sul~\ta\ej\na\cartouche{\en\ki\tu}}

\drie
\drie
 

\cartouche{\en\ki\tu}\si\po\mot~\tu\po\lan\lo\jel~~ \\ \ut\es\set\on\lum\comma \an\lan\~\lo\ten\luf
 
{\it [ENKITU].SI.PO.MOT PA.TU.PO.LAN.LO.JEL \\UT.ES.SET.ON.LUM, AN LAN  LO.TEN.LUF}

\{ENKITU-predicate-process-death\} \{past-verb-process-speak-to-God\} \{instrument-this-word-plural-last\}, \{and\} \{speak\} \{to-acquaintance-love-of-him\} 

With these last words the dying Enkidu did pray, and say to his beloved companion:  

\drie 
 
\et\Atlantan\sul~ \som\on~\an~\tem~\tu\Atlanni\noj~\\ \ek\san\pak\pik~ ~\po\ja~\su~\ej\am\si\ip~\pa\tuf\li\jun 

{\it ET.TAN.SUL.OM SOM.ON AN TEM TU.NI.NOJ \\EK.SAN.PAK.PIK PO.JA EJ.AM.SI.IP PA.TUF.LI.JUN }

\{location-dream-plural-hallucination\} \{sky-plural\} \{and\} \{earth\} \{verb-perfective shout\} \{accusative-sound\idx{sound}-bad-big\} \{progressive time\} \{person-first-predicate-one\} \{past stand direction descruction\}

"In dreams last night the heavens and the earth poured out great groans while I alone stood facing devastation.  

\drie 

\nik\si\tut\an\men\tef ~ \tu\pa\ful\et\ot ~ \li\ej\am ~ \Atlani\am\\ ~ \at\tu\pa\pus\ek\ej\am ~ \ut\nek\tak\on\ta\aj\\ ~ \lo\nap\mol\tef\ta\mot \Atlanin\es\na\cartouche{\il\kal\la} ~\\\comma\Atlanfi\tes\ta\Atlanne\las ~ \comma\aj\tu\jon

{\it NIK.SI.TUT.AN.MEN.TEF TU.PA.FUL.ET.OT LI.EJ.AM I.AM\\ AJ.TU.PA.PUS.EK.EJ.AM UT.NEK.TAK.ON.TA.AJ\\ LO.NAP.MOL.TEF.TA.MOT IN.ES.NA[IL.KAL.LA] ,\\FI.TES.TA.NE.LAS AJ.TU.JON }

\{creature predicate strong and make fear\} \{verb past fly coordinate under\} \{direction 1ST\} \{rel.clause and\} \{3rd-verb past push acc. person 1st\}\{Instrument nail sharp plural genitive 3rd\} \{direction\} \{house full fear genitive death\} \{in this [name] I.KA.LA.\} \{feminine king genitive not-light\} \{3SG verb control\} 

A fierce and threatening creature flew down at me and pushed me with its talons toward the horror-filled house of death wherein lrkalla, queen of shades, stands in command.  
 
 
\drie 

\es\aj ~ \si ~ \Atlanne\la ~ \es ~ \mul~\ek\Atlanne\so\ej~\tu\sik~\las\ta\jan\kern-3em\period

{\it LU.AJ SI NE.LAS ES MUL EK.NE.SO.EJ TU.SIK LAS.TA.JAN}

 \{place 3SG\} \{is\} \{negation light\} \{demonstrative\} \{allow\}\{acc. not some person\}\{relative clause verb see\}\{acc. light genitive day\} 

There is darkness which lets no person again see light of day.  

\drie 
 

\lu\aj ~ \si ~ \lol~ \Atlanne\li\fin\las\an\sus ~ \pu ~ \las ~ \an ~ \sus ~ \fin

{\it LU.AJ SI LOL NE.LI.FIN.LAS.AN.SUS PU LAS AN SUS FIN }

\{place 3SG\} \{is\}\{road not-direction life light and joy\} 

There is a road leading away from bright and lively life.  
 
\drie 

\lu\aj ~ \tu\puj\fin ~ \es\ej\aj\on ~  \tu\kos\toj ~ \ek\tul\suk ~ \an \\~ \Atlanne\tu\ta ~ \net\mep\tep\je ~ \lo\fan\jit\pak\ta\aj\on 

{\it LU.AJ TU.PUJ.FIN ES.EJ.AJ.ON  TU.KOS.TOJ EK.TUL.SUK AN \\NE.TU.TA NET.MEJ.TEP.JE LO.FAN.JIT.PAK.TA.AJ.ON}

\{place 3SG\} \{verb stay-alive\} \{demonstrative person 3rd plural\}\{VERB consume solid\}\{acc. dust-dry\} \{and\} \{negation verb genitive\} \{water change temperature negative\} \{dative want liquid bad genitive 3rd-plural\} 

There dwell those who eat dry dust and have no cooling water for their awful thirst.  
 
\drie

\et\ja ~ \ej\am\pa\tu\tuf ~ \et\lu\aj ~ \Atlani\ej\aj\pa\tu\sik \\~ \ek\at\es\ej\aj\on ~  \tu\pa\mot ~ \an ~ \Atlana\ek\tes\on ~ \\ ~  \mi\aj ~ \es\ej\aj ~ \pa\tu\mej\Atlanne\las ~  \\ \Atlanne\tu\ta \ek\jej\fa\an\pop\ta\aj\on 

{\it ET.JA EJ.AM.PA.TU.TUF ET.LU.AJ I.EJ.AJ.PA.TU.SIK \\EK.AT.ES.EJ.AJ.ON TU.PA.MOT AN A.EK.TES.ON \\ MI.AJ ES.EJ.AJ PA.TU.MEJ.NE.LAS \\NE.TU.TA EK.JEJ.FA.AN.POP.TA.AJ.ON }


\{coordinate time\} \{person 1st do past stand\} \{coordinate place 3rd\} \{relative clause person first verb past see\} \{acc. all demonstrative person 3rd plural\} \{verb past death\} \{and\} \{stress acc. king plural\} \{king plural\} \{in-between 3SG\} \{demonstrative 3SG person\} \{passive verb change not light\}\{acc. success far and previous genitive 3rd plural\}

As I stood there I saw all those who've died and even kings among those darkened souls have none of their remote and former glory. 
 
\drie 

\at ~ \pot\pik\ta\tem ~ \tu\pa\si\Atlanpi\Atlanem\fun ~ \Atlani\an ~ \ej\am ~ \\\pa\tu\mef\li\Atlanin~ \\\nap\ta\mot

{\it AT POT.PIK.TA.TEM TU.PA.SI.PI.EM.FUN I.AN EJ.AM \\PA.TU.MEF.LI.IN NAP.TA.MOT }


\{all\} \{good big genitive earth\} \{verb past is passive inverse find\} \{relative clause and\} \{past verb walk direction inside\} \{house genitive death\} 

All earthly greatness was forfeit and I entered then into the house of death.  
 
 
\drie


\of\on ~ \es\ej\aj ~ \pa\si\lu\aj ~ \ja\lak ~ \pa\tu\mef\li\ef~\\\Atlani\uf~\tu\mal\pot~\ek\ej\am 

{\it OF.ON ES.EJ.AJ PA.SI.LU.AJ JA.LAK PA.TU.MEF.LI.EF \\I.UF TU.MAL.POT EK.EJ.AM} 


\{other plural\} \{demonstrative 3SG person\} \{past be place 3rd\} \{time long\} \{past do come to above rel.clause intention verb greet good acc. person 1SG\}

Others who’ve been there long did rise to welcome me." 


}

