
Special thanks to, 

\noindent {\bf Ana Bosnic}¸ for her amazing guidance throughout this whole book. Ana, we wouldn’t know where would be without you. 

We vividly remember how badly we wanted you to be our supervisor. Luckily, your enthusiasm for this project was almost bigger than ours! You always had great ideas and was ready to help. We cannot thank you enough for all the proofreading you did and your many (!) comments. Your constructive criticism and thoughtful suggestions brought the quality of our book to a whole new level. Your insights and expertise added depth and credibility to our work, and we couldn’t be more grateful.  

Also, thank you for putting up with our chaotic meetings, late notifications, and last-minute room changes. You never gave up on us: even if there had to be three separate meetings to let all of use view one documentary. Your commitment and perseverance made all the difference, and we are sincerely grateful for your support. 

Ana, we had a total blast these couple of weeks! Since you will not give this course anymore, we surely hope it was just as much fun for you! Thank you a million times over! 

  

\noindent {\bf Angela Carpenter}, for her support and contribution to this project. Angela, your help and assistance have truly made an impact on this product.  

That afternoon when you were ready to lend an ear, despite the crazy time difference, was a true joy for us. How cool it was that we were talking with the professor we saw in a documentary only a few weeks before! Your patience and genuine interest in our non-existing language made our brainstorming session truly delightful. We owe you a debt of gratitude for sharing all that knowledge from your classes. Your enthusiasm, and encouragement meant the world to us! 

 

\noindent {\bf Humanities Honours Programme}, for funding this project,  and basically making a dream come true. Your help was like finding a hidden treasure in the bustling streets of Hong Kong. Also thank you for providing the coffee to keep us going. Your contribution will forever be cherished and remembered.
